---
title: "Fix DADA2 Errors Related to Loading Rcpp in R"
author: "Thieu Nguyen"
date: "2024-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dada2 is a plugin from QIIME2 needed for targeted gene metagenomic data analysis (16S sequencing). I have errors when running dada2. The error messeage is:

## "An error was encountered while running DADA2 in R"

I want to run these codes

```{r eval=FALSE, include=TRUE}
system("qiime dada2 denoise-single \
  --i-demultiplexed-seqs demux.qza \
  --p-trim-left 13 \
  --p-trunc-len 150 \
  --o-representative-sequences asv-sequences.qza \
  --o-table asv-table.qza \
  --o-denoising-stats denoising-stats.qza \
  --verbose")
```

However, an error happens:

R version 4.4.2 (2024-10-31) 

Loading required package: Rcpp

Error: package or namespace load failed for ‘Rcpp’ in library.dynam(lib, package, package.lib): shared object ‘Rcpp.so’ not found

Error: package ‘Rcpp’ could not be loaded

.....

Plugin error from dada2:

An error was encountered while running DADA2 in R (return code 1), please inspect stdout and stderr to learn more.

I check errors in Rstudio but everything is OK. So the errors come from loading R in Terminal.

## Fixing errors

The error indicates that 'Rcpp' and its dependencies are not installed properly. I need to install Rcpp's dependencies one by one as follow: 

The place I install dada2 is in the Terminal of Macbook, not in RStudio.

Open Terminal and activate 'qiime2-amplicon-2024.10'

NOTE: DO NOT TYPE 'system(" ")' IN TERMINAL, just type the codes in double quotes " " of the phrase 'system(" ")'. For examples:

conda activate qiime2-amplicon-2024.10

```{r eval=FALSE, include=TRUE}
system("conda activate qiime2-amplicon-2024.10")
```

Then open R in 'qiime2-amplicon-2024.10' by typing 'R'

```{r eval=FALSE, include=TRUE}
system("R")
```

Now, in R in (qiime2-amplicon-2024.10) nnthieu@192 data %,
I install dependencies needed.


```{r eval=FALSE, include=TRUE}
BiocManager::install("Rcpp", force=TRUE)

# then check dada2 loading
library(dada2)
```

I face another error related to other dependencies like 'reshape2.so not found'.
I continue to install dependencies as above


```{r eval=FALSE, include=TRUE}
BiocManager::install("reshape2", force=TRUE)

# then check dada2 loading
library(dada2)
```

Until dada2 can be loaded normally as


> library(dada2)

Loading required package: Rcpp


OK, it is good now.

## Run dada2 again


```{r eval=FALSE, include=TRUE}
system("qiime dada2 denoise-single \
  --i-demultiplexed-seqs demux.qza \
  --p-trim-left 13 \
  --p-trunc-len 150 \
  --o-representative-sequences asv-sequences.qza \
  --o-table asv-table.qza \
  --o-denoising-stats denoising-stats.qza ")
```

Now dada2 works properly.

-------

